|

### 2023-05-11 paper/00-main.Rmd
fixed effects table
abstract

# abstract from BDB submission form
# rewrite this
abstract: |
  This paper evaluates pass rusher performance in the National Football League. We propose STRAIN, a new measure for pass rusher effectiveness, which accounts for the distance between the pass rusher and quarterback and the rate at which this distance is being reduced. STRAIN is a simple and interpretable statistic, with higher STRAIN corresponds to greater pass rushing ability. Our metric is closely related to the concept of "strain rate" in materials science. We present the mathematical motivations for our metric, followed by an illustration with an example play. The analysis concludes with a comparison of STRAIN for different positions and our rankings of NFL's best pass-rushers according to our metric.

bibliography: bibliography.bib
biblio-style: chicago
output: rticles::asa_article
# header-includes:
#   \renewcommand{\arraystretch}{0.8}
---

>>>>

abstract: |
  In American football, existing metrics for assessing pass rush are either discrete-time quantities or based on subjective judgement. Using player tracking data, we propose STRAIN, a novel metric for evaluating pass rushers in the National Football League (NFL) at the continuous-time within-play level. Inspired by the concept of strain rate in materials science, STRAIN is a simple and interpretable means for measuring defensive pressure in football. It is a directly-observed quantity as a function of two features: the distance between the pass rusher and quarterback, and the rate at which this distance is being reduced. Our metric possesses great predictability of pressure and stability over time. We use a multilevel model to isolate the contribution of every pass rusher in a play while controlling for their pass blocker matchup, the quarterback, and play-level information. We apply our approach to NFL data and present results for the first 8 weeks of the 2021 season. These include comparisons of STRAIN for different positions and play outcomes, and rankings of the NFL's best pass rushers according to our metric.

bibliography: bibliography.bib
biblio-style: chicago
output: rticles::asa_article
header-includes:
  \usepackage{mathptmx}
  # \renewcommand{\arraystretch}{0.8}
---

### 2023-05-11 paper/01-introduction.Rmd
abstract

The first BDB led to several contributions largely focused on offensive performance evaluation. For example, one group of finalists introduced an approach for modeling the hypothetical completion probability of a pass aiding in the evaluation of quarterback (QB) decision making [@deshpande2020expected]. The winners of the inaugural BDB focused on identifying receiver routes via clustering techniques [@chu2020route] and convolutional neural networks [@sterken2019routenet]. Along with the BDB competition entries, the public release of NGS data allowed researchers to tackle (pun intended) a variety of other problems such as revisiting fourth down decision making [@lopez2020bigger], annotating pass coverage with Gaussian mixture models [@dutta2020unsupervised], and introducing a continuous-time framework to estimate within-play value [@yurko2020going].  Since its inception, the BDB has chosen a different theme each year leading to new insight about evaluating different positions such as running backs, defensive backs, and special teams. The [2023 edition of the NFL BDB](https://www.kaggle.com/competitions/nfl-big-data-bowl-2023) asked participants to evaluate lineman on passing plays [@Howard2023NFL].

@@

<!-- Previous research bdb -->
<!-- this year: linemen -->
<!-- Importance of D-linemen, pass rush (Maybe talk about role of each position, edge vs interior, defensive formation) -->

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before the QB is able to throw a pass \\
Hits & A defender tackles the QB behind the line of scrimmage after the QB is able to throw a pass. \\
Hurries & A defender pressures the QB behind the line of scrimmage such that the QB must release the ball prior sooner than they wanted to. \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
\hline
\end{tabular}
\end{table}

>>>>

The first BDB led to several contributions largely focused on offensive performance evaluation. For example, one group of finalists introduced an approach for modeling the hypothetical completion probability of a pass aiding in the evaluation of quarterback (QB) decision making [@deshpande2020expected]. The winners of the inaugural BDB focused on identifying receiver routes via clustering techniques [@chu2020route] and convolutional neural networks [@sterken2019routenet]. Along with the BDB competition entries, the public release of NGS data allowed researchers to tackle (pun intended) a variety of other problems such as revisiting fourth down decision making [@lopez2020bigger], annotating pass coverage with Gaussian mixture models [@dutta2020unsupervised], and introducing a continuous-time framework to estimate within-play value [@yurko2020going].  Since its inception, the BDB has chosen a different theme each year leading to new insight about evaluating different positions such as running backs, defensive backs, and special teams. The 2023 edition of the NFL Big Data Bowl asked participants to evaluate linemen on passing plays [@Howard2023NFL].

## Linemen in American Football

<!-- Importance of D-linemen, pass rush (Maybe talk about role of each position, edge vs interior, defensive formation) -->

@@

@@

\begin{table}
\caption{A summary of previously-existing pass rush metrics. \label{tab:metrics}}
\centering
\begin{tabular}{p{0.27\textwidth}p{0.67\textwidth}}
\hline
Metric & Description \\ 
\hline
Sacks & A defender tackles the QB behind the line of scrimmage before the QB is able to throw a pass \\
Hits & A defender tackles the QB behind the line of scrimmage after the QB is able to throw a pass. \\
Hurries & A defender pressures the QB behind the line of scrimmage forcing the QB to throw the ball prior sooner than intended. \\
Pressures & $\text{Hurries} + \text{ Hits } + \text{ Sacks}$ \\
Pass-Rush Productivity & $\displaystyle \frac{(\text{Hurries} + \text{Hits})/2 + \text{Sacks}}{\text{ Pass Rush Snaps}}$\\
Time In Pocket & Time (in seconds) between ball snap and throw or pocket collapse for a QB \\
Pass Rush Win Rate & Rate at which pass rusher beats pass block within 2.5 seconds after ball snap\\
\hline
\end{tabular}
\end{table}

### 2023-05-11 paper/03-methods.Rmd
abstract

## Application to pass rush in football

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (in terms of elapsed time between ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

>>>>

## Application to pass rush in football {#sec:app}

@@

We consider a normal distribution to obtain the random intercepts for each player toward their respective position group means.
This provides us with average individual effects on STRAIN for each player while applying shrinkage, which is a useful property in this problem's context since we do not observe the same number of plays for each player.
With regards to the fixed-effects about pass rusher $j$ in play $i$, we include the length of the play (i.e. the elapsed time between the ball snap and a QB event), the positions of pass rusher and nearest blocker, and the number of pass blockers assigned to the pass rusher based on the PFF data.
Table \ref{tab:pos} shows our positional categorization for the pass rush and pass block roles.
We implement the model using penalized likelihood via the `lme4` package in `R` [@bateslme4; @R2023Language].

### 2023-05-11 paper/04-results.Rmd
abstract

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass-rushers (NT's and DT's) separately from edge rushers (OLB's and DE's).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks for each defender.

@@

```{r}
#| label: fig_rankings_boot
#| fig.cap: "Top 10 pass rushers for each position (DE, DT, NT, OLB) by mean varying intercept."
#| fig.width: 6
#| fig.height: 8
source(here("scripts", "10_multilevel_mod.R"))
fig_rankings_boot
```

>>>>

Since STRAIN is observed continuously for every play in our data, this allows us to aggregate across all frames played and compute the average STRAIN for NFL pass rushers over the course of the eight-week sample size, as discussed in Section \ref{sec:app}.
Based on the clearly distinct patterns for different positions as previously observed, we evaluate interior pass-rushers (NT's and DT's) separately from edge rushers (OLB's and DE's).
Tables \ref{tab:edge} and \ref{tab:interior} are leaderboards for the NFL's best edge and interior rushers (with at least 100 plays) rated by the average STRAIN across all frames for the first eight weeks of the 2021 regular season.
The tables also consist of the total number of hits, hurries, and sacks for each defender.

@@

```{r}
#| label: fig_rankings_boot
#| fig.cap: "Top 10 pass rushers for each position (DE, DT, NT, OLB) by mean varying intercept with 95% confidence intervals."
#| fig.width: 6
#| fig.height: 8
source(here("scripts", "10_multilevel_mod.R"))
fig_rankings_boot
```
